Pulmonary acini exhibit complex changes during postnatal rat lung development

### 2021-06-03 manuscript.tex
Fix typos [ci skip]
This build is based on
https://github.com/habi/acinar-analysis-manuscript/commit/faa233464aaa9be847c35036707b917ec510a8af.

This commit was created by the following CI build and job:
https://github.com/habi/acinar-analysis-manuscript/commit/faa233464aaa9be847c35036707b917ec510a8af/checks
https://github.com/habi/acinar-analysis-manuscript/runs/902824374

A preprint of this manuscript is available on \emph{bioRxiv} with the DOI \href{https://doi.org/10.1101/786681}{10.1101/786681}.
It has been submited to \href{https://journals.plos.org/plosone/}{\emph{PLOS ONE}}.

This manuscript
(\href{https://habi.github.io/acinar-analysis-manuscript/v/457b81bb713828ba723145280d0ee7c4a8b18609/}{permalink})
was automatically generated
from \href{https://github.com/habi/acinar-analysis-manuscript/tree/457b81bb713828ba723145280d0ee7c4a8b18609}{habi/acinar-analysis-manuscript@457b81b}
on April 9, 2021.

@@

\begin{itemize}
\item
  \textbf{David Haberthür}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3388-9187}{0000-0003-3388-9187}
  · \href{mailto:haberthuer@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(457b81b)}{haberthuer@ana.unibe.ch}
  · \includegraphics{images/github.svg}
  \href{https://github.com/habi}{habi}
  · \includegraphics{images/twitter.svg}
  \href{https://twitter.com/habi}{habi}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Eveline Yao}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Sébastien F. Barré}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Tiziana P. Cremona}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-0061-9578}{0000-0003-0061-9578}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Stefan A. Tschanz}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3880-4437}{0000-0003-3880-4437}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Johannes C. Schittny}✉
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-4025-3961}{0000-0003-4025-3961}
  · \href{mailto:schittny@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(457b81b)}{schittny@ana.unibe.ch}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\end{itemize}

@@

As shown in a study by Osmanagic et al.~{[}\protect\hyperlink{ref-LzalnLBH}{49}{]} the total alveolar number shows a large variability between different labs.
E.g.~the reported total number of alveoli in adult C57BL6 mice differs by a factor of approximately 8.
All the studies mentioned my Osmanagic et al.~were performed in labs well-known for their quality of stereology-based investigations of lung morphology.
Therefore, any simple bias or methodical error may not account for the reported differences.
It seems to be that the total number of alveoli is dependent on unknown small methodical differences and minute differences in how the operator actually does count.
The seemingly minuscule volume fraction of lung studied (around 1 \% or less) in this manuscript is---according to the stereological guidelines {[}\protect\hyperlink{ref-dNc8FfNn}{28}{]}---ample volume to gain mathematically correct results if the stereology is carried out correctly as we have done to the best of our knowledge.
For human lungs for example a so-called sampling cascade is used to study details down to the thickness of the air-blood barrier (0.5 µm, see Figure 1 of {[}\protect\hyperlink{ref-1CLJ32t5m}{50}{]} for a visualization of this cascade).
The studied volume in this case is several decimal powers less than what is studied here and still mathematically sound and accepted.

@@

Multi-breath gas-washout is also dependent on the size of the acini.
Based on computational simulations of nitrogen multi-breath gas-washout it has been shown that small acini washout faster at the beginning and slower at the end of the washout as compared to larger acini {[}\protect\hyperlink{ref-eb0gk6VO}{13}{]}.
Not surprisingly, the distribution of the acinar sizes in terms of their range has also an influence on the gas-washout.
Furthermore, a clustering versus an even distribution of the locations of acinar similar sizes effects gas-washout.
Therefore, the question where an acinus of which size is located in the lungs has a significant influence on pulmonary ventilation and particle deposition.

@@

@@

\leavevmode\hypertarget{ref-k5fqLjiy}{}%
54. Sznitman J, Heimsch F, Altorfer D, Schittny JC, Rösgen T. Alveolar flow simulations during rhythmical breathing motion in reconstructed XTM acinar airspaces. Journal of Biomechanics. 2006;39: S441. doi:\href{https://doi.org/10.1016/s0021-9290(06)84801-3}{10.1016/s0021-9290(06)84801-3}
\end{cslreferences}

>>>>

A preprint of this manuscript is available on \emph{bioRxiv} with the DOI \href{https://doi.org/10.1101/786681}{10.1101/786681}.
It has been submited to \href{https://journals.plos.org/plosone/}{\emph{PLOS ONE}}.
The manuscript \href{https://github.com/habi/acinar-analysis-manuscript/compare/5e778036e316c47b8bac4ce50086b191a7e3707a...faa233464aaa9be847c35036707b917ec510a8af}{has been updated} since submission.

This manuscript
(\href{https://habi.github.io/acinar-analysis-manuscript/v/faa233464aaa9be847c35036707b917ec510a8af/}{permalink})
was automatically generated
from \href{https://github.com/habi/acinar-analysis-manuscript/tree/faa233464aaa9be847c35036707b917ec510a8af}{habi/acinar-analysis-manuscript@faa2334}
on June 3, 2021.

@@

\begin{itemize}
\item
  \textbf{David Haberthür}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3388-9187}{0000-0003-3388-9187}
  · \href{mailto:haberthuer@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(faa2334)}{haberthuer@ana.unibe.ch}
  · \includegraphics{images/github.svg}
  \href{https://github.com/habi}{habi}
  · \includegraphics{images/twitter.svg}
  \href{https://twitter.com/habi}{habi}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Eveline Yao}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Sébastien F. Barré}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Tiziana P. Cremona}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-0061-9578}{0000-0003-0061-9578}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Stefan A. Tschanz}
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-3880-4437}{0000-0003-3880-4437}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\item
  \textbf{Johannes C. Schittny}✉
  · \includegraphics{images/orcid.svg}
  \href{https://orcid.org/0000-0003-4025-3961}{0000-0003-4025-3961}
  · \href{mailto:schittny@ana.unibe.ch?subject=Feedback\%20regarding\%20the\%20acinus\%20manuscript\%20(faa2334)}{schittny@ana.unibe.ch}
  Institute of Anatomy, University of Bern, Bern, Switzerland
\end{itemize}

@@

As shown in a study by Osmanagic et al.~{[}\protect\hyperlink{ref-LzalnLBH}{49}{]} the total alveolar number shows a large variability between different labs.
E.g.~the reported total number of alveoli in adult C57BL6 mice differs by a factor of approximately 8.
All the studies mentioned by Osmanagic et al.~were performed in labs well-known for their quality of stereology-based investigations of lung morphology.
Therefore, any simple bias or methodical error may not account for the reported differences.
It seems to be that the total number of alveoli is dependent on unknown small methodical differences and minute differences in how the operator actually does count.
The seemingly minuscule volume fraction of lung studied (around 1 \% or less) in this manuscript is---according to the stereological guidelines {[}\protect\hyperlink{ref-dNc8FfNn}{28}{]}---ample volume to gain mathematically correct results if the stereology is carried out correctly as we have done to the best of our knowledge.
For human lungs for example a so-called sampling cascade is used to study details down to the thickness of the air-blood barrier (0.5 µm, see Figure 1 of {[}\protect\hyperlink{ref-1CLJ32t5m}{50}{]} for a visualization of this cascade).
The studied volume in this case is several decimal powers less than what is studied here and still mathematically sound and accepted.

@@

Multi-breath gas-washout is also dependent on the size of the acini.
Based on computational simulations of nitrogen multi-breath gas-washout it has been shown that small acini washout faster at the beginning and slower at the end of the washout as compared to larger acini {[}\protect\hyperlink{ref-eb0gk6VO}{13}{]}.
Not surprisingly, the distribution of the acinar sizes also also an influence on the gas-washout.
Furthermore, a clustering versus an even distribution of the locations of similar acinar sizes also affects gas-washout.
Therefore, the question where an acinus of which size is located in the lungs has a significant influence on pulmonary ventilation and particle deposition.

@@

We thank \texttt{manubot} {[}\protect\hyperlink{ref-YuJbg3zO}{55}{]} for providing us with the means to collaboratively write and craft this manuscript.

@@

\leavevmode\hypertarget{ref-k5fqLjiy}{}%
54. Sznitman J, Heimsch F, Altorfer D, Schittny JC, Rösgen T. Alveolar flow simulations during rhythmical breathing motion in reconstructed XTM acinar airspaces. Journal of Biomechanics. 2006;39: S441. doi:\href{https://doi.org/10.1016/s0021-9290(06)84801-3}{10.1016/s0021-9290(06)84801-3}

\leavevmode\hypertarget{ref-YuJbg3zO}{}%
55. Himmelstein DS, Rubinetti V, Slochower DR, Hu D, Malladi VS, Greene CS, et al. Open collaborative writing with Manubot. PLOS Computational Biology. 2019;15: e1007128. doi:\href{https://doi.org/10.1371/journal.pcbi.1007128}{10.1371/journal.pcbi.1007128}
\end{cslreferences}
